---
title: "Understanding Society Codebook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding Society Codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(lfsclean)
```

## List of Variables

This document details the cleaned variables from the Understanding Society dataset which are available through using the `ukhlsclean` package, grouped by cleaning function category. Where not otherwise specified, variables are available in all waves of the data

### __Global variables__

These variables are identifier and weighting variables produced automatically and should not be specified in the `keep_vars` argument to the `ukhlsclean` function. 

- __id__ (numeric) individual identifier. 
    - Constructed from the `pidp` variable for individuals sampled by the UKHLS or from the `pid` variable for              individuals carried over from the British Household Panel Survey (BHPS). For those individuals, __id__ can be         linked to __pid__ in the BHPS data.
    
- __hidp__ (numeric) household identifier. 

- __wave_no__ (numeric) wave the observation is taken from.

- __bhps_sample__ (logical) if TRUE, the individual was carried over from the BHPS. 

- __nwaves__ (numeric) number of waves the individual has observations for. 

- __wave_(1-11)__ (numeric) individual identifier, equal to 1 if the individual was observed in the respective wave, 0 otherwise.

- __year__ (numeric) interview date - year.

- __month__ (numeric) interview date - month.

- __day__ (numeric) interview date - day.

- __weight_xw__ (numeric) cross-sectional survey weight, for weighting one wave to the population.

### __Demographic variables__ 

Variables produced by the demographic module `ukhls_clean_demographic`. Variables in this category are prefixed with a `d_`. 

- __d_age__ (numeric) continuous, single-year of age

- __d_age_5cat__ (factor) age category - 5 categories:
    -  "16-24"
    -  "25-34"
    -  "35-49"
    -  "50-64"
    -  "65+"

- __d_age_12cat__ (factor) age category - 12 categories:
    -  "16-19"
    -  "20-24"
    -  "25-29"
    -  "30-34"
    -  "35-39"
    -  "40-44"
    -  "45-49"
    -  "50-54"
    -  "55-59"
    -  "60-64"
    -  "65-69"
    -  "70+"

- __d_sex__  (factor) sex:
  - "male"
  - "female"

- __d_gor__  (factor) government office region:
  - "north_east"
  - "north_west"
  - "yorkshire"
  - "east_midlands"
  - "west_idlands"
  - "east"
  - "london"
  - "south_east"
  - "south_west"
  - "wales"
  - "scotland"
  - "northern_ireland"
  
- __d_area__ (factor) urban or rural area:
  - "urban"
  - "rural"
  
- __d_marstat__ (factor) marital status:
  - "single"
  - "married"
  - "sep_div_wid" (separated, divorced, or widowed)
  
- __d_hiqual__ (factor) highest qualification level achieved:
  - "degree"
  - "other_he"
  - "alevel" 
  - "gcse"
  - "other_qual"
  - "no_qual"
  
- __d_ethnicity_2cat__  (factor) ethnicity - 2 categories:
  - "white"
  - "non_white"

- __d_ethnicity_5cat__  (factor) ethnicity - 5 categories:
   - "white"
   - "black"
   - "asian"
   - "mixed"
   - "other"
   
- __d_ethnicity_9cat__  (factor) ethnicity - 9 categories:
   - "white_british"
   - "white_non_british"
   - "mixed"
   - "indian"
   - "pakistani"
   - "bangladeshi"
   - "other_asian"
   - "black"
   - "other"
   
### __Household variables__

Variables produced by the household module `ukhls_clean_hhold`. Variables in this category are prefixed with a `hh_`. 

- __hh_hometenure__ (factor) household occupancy type:
   - "owner_occupier"
   - "private_renter"
   - "social_renter"
   
- __hh_age_yngchl__ (factor) age of the youngest child in the household:
   - "no_children"
   - "0-2"
   - "3-4"
   - "5-11"
   - "12-15"
   
- __hh_type_6cat__ (factor) household composition:
   - "1 adult 0 children"
   - "1 adult 1+ children"
   - "2 adult 0 children"
   - "2 adult 1+ children"
   - "3+ adult 0 children"
   - "3+ adult 1+ children"
   
- __hh_numadult__ (numeric) number of adults in the household 

- __hh_numchild__ (numeric) number of children in the household 

- __hh_size__ (numeric) household size (number of adults plus children) 
   
### __Economic status variables__

Variables produced by the economic status module `ukhls_clean_econstat`. Variables in this category are prefixed with an `l_`
   
- __l_econ_stat_2cat__ (factor) economic status, 2 category:
  - "employed"
  - "not_employed"
  
- __l_econ_stat_3cat__ (factor) economic status, 3 category:
  - "employed"
  - "unemployed"
  - "inactive"
  
- __l_econ_stat_7cat__ (factor) economic status, 7 category:
  - "employed"
  - "self_employed"
  - "unemployed"
  - "sick"
  - "retired"
  - "education"
  - "other"
  
- __l_grss_earnings_usual__ (numeric) nominal gross earnings per month from employment (usual pay) plus self-employment

- __l_grss_earnings_last__ (numeric) nominal gross earnings per month from employment (last received payment) plus self-employment

- __l_real_grss_earnings_usual__ (numeric) real gross earnings per month from employment (usual pay) plus self-employment

- __l_real_grss_earnings_last__ (numeric) real gross earnings per month from employment (last received payment) plus self-employment

   
### __Health and wellbeing variables__ 

Variables produced by the health and wellbeing module `ukhls_clean_health`. Variables in this category are prefixed with a `h_`. 

- __h_eq5d_score__  (numeric) EQ-5D utility score:
   - health utility mapped from the SF-12 variables in Understanding Society using the [mapping algorithm](https://www.herc.ox.ac.uk/downloads/downloads-supporting-material-1/sf-12-responses-and-eq-5d-utility-values) developed by:
   - Gray A, Rivero-Arias O, and Clarke P (2006). Estimating the association between SF-12 responses and EQ-5D utility values by response mapping. *Medical Decision Making*; 26(1):18-29
   
- __h_pregnant__ (factor) currently pregnant:
  - "not_pregnant"
  - "pregnant"


- __h_disability__  (factor) individual has a disability:
   - "disability"
   - "no_disability"


- __h_care_hhold__  (factor) individual cares for someone in their household:
   - "carer"
   - "not_carer"
   
- __h_satisfaction_health__  (factor) satisfaction with health:
   - "completely_disatisfied"
   - "mostly_disatisfied"
   - "somewhat_disatisfied"
   - "neutral"
   - "somewhat_satisfied"
   - "mostly_satisfied"
   - "completely_satisfied"
   
- __h_satisfaction_life__  (factor) satisfaction with life overall:
   - "completely_disatisfied"
   - "mostly_disatisfied"
   - "somewhat_disatisfied"
   - "neutral"
   - "somewhat_satisfied"
   - "mostly_satisfied"
   - "completely_satisfied"

### __Alcohol variables__ 

Variables produced by the alcohol and drinking module `ukhls_clean_alcohol`. Variables in this category are prefixed with a `a_`. 

- __a_current_abstainer__  (factor) currently an abstainer from drinking alcohol (past 12 months) __(waves 7, 9, 11)__:
  - "no"
  - "yes"

- __a_always_abstainer__  (factor) has always been an abstainer from drinking alcohol __(waves 7, 9, 11)__:
  - "no"
  - "yes"

- __a_ndrinks__  (factor) number of drinks in a typical day when drinking __(waves 7, 9, 11)__:
  - "0"
  - "1-2"
  - "3-4"
  - "5-6"
  - "7-9"
  - "10+"
  
- __a_freq_binge__  (factor) frequency binge drinks. Frequency of drinking more than 6 (female) or 8 (male) units on a single occassion in the last year __(waves 7, 9, 11)__:
  - "Never"
  - "Less than monthly"
  - "Monthly"
  - "Weekly"
  - "Daily"
