---
title: "Understanding Society Codebook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding Society Codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(lfsclean)
```

## List of Variables

This document details the cleaned variables from the Understanding Society dataset which are available through using the `ukhlsclean` package, grouped by cleaning function category.

### __Global variables__

These variables are identifier and weighting variables produced automatically and should not be specified in the `keep_vars` argument to the `ukhlsclean` function. 

- __id__ (numeric) individual identifier. 
    - Constructed from the `pidp` variable for individuals sampled by the UKHLS or from the `pid` variable for              individuals carried over from the British Household Panel Survey (BHPS). For those individuals, __id__ can be         linked to __pid__ in the BHPS data.
    
- __hidp__ (numeric) household identifier. 

- __wave_no__ (numeric) wave the observation is taken from.

- __bhps_sample__ (logical) if TRUE, the individual was carried over from the BHPS. 

- __nwaves__ (numeric) number of waves the individual has observations for. 

- __wave_(1-11)__ (numeric) individual identifier, equal to 1 if the individual was observed in the respective wave, 0 otherwise.

- __year__ (numeric) interview date - year.

- __month__ (numeric) interview date - month.

- __day__ (numeric) interview date - day.

- __weight_xw__ (numeric) cross-sectional survey weight, for weighting one wave to the population.

### __Demographic variables__ 

Variables produced by the demographic module `ukhls_clean_demographic`. Variables in this category are prefixed with a `d_`. 

- __d_age__ (numeric) continuous, single-year of age

- __d_age_5cat__ (factor) age category - 5 categories:
    -  "16-24"
    -  "25-34"
    -  "35-49"
    -  "50-64"
    -  "65+"

- __d_age_12cat__ (factor) age category - 12 categories:
    -  "16-19"
    -  "20-24"
    -  "25-29"
    -  "30-34"
    -  "35-39"
    -  "40-44"
    -  "45-49"
    -  "50-54"
    -  "55-59"
    -  "60-64"
    -  "65-69"
    -  "70+"

- __d_sex__  (factor) sex:
  - "male"
  - "female"

- __d_gor__  (factor) government office region:
  - "north_east"
  - "north_west"
  - "yorkshire"
  - "east_midlands"
  - "west_idlands"
  - "east"
  - "london"
  - "south_east"
  - "south_west"
  - "wales"
  - "scotland"
  - "northern_ireland"
  
- __d_area__ (factor) urban or rural area:
  - "urban"
  - "rural"
  
- __d_marstat__ (factor) marital status:
  - "single"
  - "married"
  - "sep_div_wid" (separated, divorced, or widowed)
  
- __d_hiqual__ (factor) highest qualification level achieved:
  - "degree"
  - "other_he"
  - "alevel" 
  - "gcse"
  - "other_qual"
  - "no_qual"
  
- __d_ethnicity_2cat__  (factor) ethnicity - 2 categories:
  - "white"
  - "non_white"

- __d_ethnicity_5cat__  (factor) ethnicity - 5 categories:
   - "white"
   - "black"
   - "asian"
   - "mixed"
   - "other"
   
- __d_ethnicity_9cat__  (factor) ethnicity - 9 categories:
   - "white_british"
   - "white_non_british"
   - "mixed"
   - "indian"
   - "pakistani"
   - "bangladeshi"
   - "other_asian"
   - "black"
   - "other"
   
   
### __Health and wellbeing variables__ 

